---
title: "ReproductiveResearch1"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=TRUE}
library(dplyr)
library(plyr)
library(ggplot2)

```

### Code for reading in the dataset and/or processing the data

```{r reading, include=TRUE, echo = TRUE}
if(!file.exists("getdata-projectfiles-UCI HAR Dataset.zip")) {
  temp <- tempfile()
  download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp)
  unzip(temp)
  unlink(temp)
}

data <- read.csv("activity.csv")

```

### Histogram of the total number of steps taken each day

```{r hist, include=TRUE, echo = TRUE}

steps_by_day <- aggregate(steps ~ date, data, sum)
hist(steps_by_day$steps, main = "Total Steps Each Day", col = "aquamarine", xlab = "Number of Steps")

```

### Mean and median number of steps taken each day

```{r meanmedian, include=TRUE, echo = TRUE}

rmean <- mean(steps_by_day$steps)
cat("The mean of the data is: ", rmean)
rmedian <- median(steps_by_day$steps)
cat("The median of the data is ", rmedian)

```

### Time series plot of the average number of steps taken

```{r avgsteps, include=TRUE, echo = TRUE}

steps_by_interval <- aggregate(steps ~ interval, data, mean)
par(bg = 'lightblue')
plot(steps_by_interval$interval, steps_by_interval$steps, type = "l", xlab = "Interval", ylab = "Number of Steps", main = "Average Number of Steps per Day by Interval")

```

### The 5-minute interval that, on average, contains the maximum number of steps

```{r max5int, include=TRUE, echo = TRUE}

max_interval <- steps_by_interval[which.max(steps_by_interval$steps),1]
cat("The maximum interval is: ", max_interval)

```

### Code to describe and show a strategy for imputing missing data

```{r stratimpu, include=TRUE, echo = TRUE}

incomplete <- sum(!complete.cases(data))
imputed_data <- transform(data, steps = ifelse(is.na(data$steps), steps_by_interval$steps[match(data$interval, steps_by_interval$interval)], data$steps))

imputed_data[as.character(imputed_data$date) == "2012-10-01", 1] <- 0

```

### Histogram of the total number of steps taken each day after missing values are imputed

```{r histsteps, include=TRUE, echo = TRUE}

steps_by_day_i <- aggregate(steps ~ date, imputed_data, sum)
hist(steps_by_day_i$steps, main = "Total Steps Each Day", col="red", xlab = "Number of Steps")

```

### Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r panelPlot, include=TRUE, echo=TRUE, echo = TRUE}

# Calculate new mean and median for imputed data
rmean.i <- mean(steps_by_day_i$steps)
rmedian.i <- median(steps_by_day_i$steps)

# Calculate difference between imputed and non-imputed data
mean_diff <- rmean.i - rmean
med_diff <- rmedian.i - rmedian

# Total difference
total_diff <- sum(steps_by_day_i$steps) - sum(steps_by_day$steps)

weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
imputed_data$dow = as.factor(ifelse(is.element(weekdays(as.Date(imputed_data$date)),weekdays), "Weekday", "Weekend"))

steps_by_interval_i <- aggregate(steps ~ interval + dow, imputed_data, mean)

library(lattice)

xyplot(steps_by_interval_i$steps ~ steps_by_interval_i$interval|steps_by_interval_i$dow, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")

```



